Efficient Planning for Sokoban with Answer Set Programming - A Case Study

### 2018-08-20 BATexFiles/Draft.tex
a little beauty correction for the tex file, had a lot of // for new lines, now replaced by \paragraph

\documentclass{article}
\title{Efficient Planning for Sokoban with Answer Set Programming - A Case Study}
\author{Thomas Verweyen}
\date{\today}
\usepackage{listings}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{color}
\lstset{numbers=left,
literate = {-}{-}1,
basicstyle=\footnotesize,
numbersep=5pt, backgroundcolor=\color{white},
captionpos=b,keywordstyle=\color{blue},
commentstyle=\color{green}}
\begin{document}
\maketitle
\renewcommand*\contentsname{Summary}
\tableofcontents
\section*{Abstract}
Sokoban approach with ASP planning $\rightarrow$ PlainEnc, IncEnc, Optimization (?)
\section{Introduction}
%%%%TODO: REWORK THIS
What do i do? I'm working on the Sokoban problem and trying to do my best to find a fast and easy solution to it. I'm not working on any level but just the ones meeting some criteria. I'm picking qualitative and quantitative characteristics and deciding on some values to examine. My encodings are supposed to work better on some of the instances and worse on others. What rules describe those fluctuations? What assumptions can be made? Is there a trend when comparing runtimes to my qualitative and quantitative characteristics? E.g. you would expect the runtime to go up when the instance's size grows or when the amount of boxes in the level grows.
In this thesis I will work on optimizing ASP encodings for the game Sokoban towards different instances. First Ill briefly explain what Sokoban is and how it is played, then Ill give a short introduction on comparable/similar/other scientific studies concerning sokoban. Then Ill go over my work, ... and finally evaluation and Discussion of Results.
\section{What is Sokoban}
\subsection{Background}
%% ZU KURZ, MIR FEHLT EIN BISSCHEN LORE (SOKOBAN HEISST WARENHAUSMANN, ENTWICKELT VON.., ...) UND EIN BISSCHEN FORMALITÄT (VGL \cite{Froleyks2017UsingAA})
Sokoban is a logistic planning puzzle game from Japan. Sokoban was apparently created by Hiroyuki Imabayashi in 1981, and published by \textit{Thinking Rabbit} in 1982. Sokoban means warehouse person in japanese. The game has very simple rules but shows to be very complex for humans and for AI. This makes Sokoban a very interesting subject for research regarding planning problems.
%LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
\subsection{Rules}
\begin{figure}[h]
\centering
\includegraphics[scale=0.2]{SokobanLevel1}
\caption{First of the Classical Sokoban Levels}
\label{fig:1}
\end{figure}
Figure \ref{fig:1} illustrates the first Sokoban level of the original set of levels. The game is played on a board that can be modelled as a two-dimensional grid. The squares can either be floor or wall squares. Some floor squares are marked as goals and the same number of floor squares contain boxes. The player pushes boxes around this modelled warehouse, trying to cover the marked goal fields (targets) with one of the boxes. For that the player can move one step at a time into one of the four directions \textit{up}, \textit{down}, \textit{left} and \textit{right}. When the player moves onto a box's position the box is pushed one field into the direction of the move. There can never be multiple boxes on one field, neither can there be a box where there is no field. A \textit{solution} to a level is a sequence of moves that leaves a box on each target.
In many of the levels created to this date, the warehouse has a lot of small passageways and very intricate designs. In this thesis I will look at more open level designs, e.g. rectangular levels.
Sokoban is a very difficult planning problem for multiple reasons. There are a lot of possibilities to create a deadlock (a gamestate from which no solution can be found), the branching factor is large and the optimal solution can be very long \cite{BoteaHeuristicsVsPlanning}.

@@

@@

\subsubsection{Incremental Encoding}
\lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
The incremental encoding is similar to the plain encoding. The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t (STEP ALSO IS THOUGH). This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable.

\subsubsection{Advanced Encoding}
\lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
\subsection{Evaluation}
\subsection{Experiments and Results}
\section{Discussion of Results}
compare chapter State of the Art with chapter My Work \cite{LifschitzASP}

\begin{thebibliography}{}
\bibitem{LifschitzASP}
Vladimir Lifschitz
\textit{Answer Set Planning}
ICLP, 1999.

\bibitem{Lifschitz2002AnswerSP}
Vladimir Lifschitz
\textit{Answer set programming and plan generation}
Artificial Intelligence, 2002, Vol. 138, p.39-54.

\bibitem{LifschitzWhatisASP}
Vladimir Lifschitz
\textit{What is Answer Set Programming?}
AAAI, 2008.

\bibitem{Gebser2013RicochetRA}
Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer
\textit{Ricochet Robots: A Transverse ASP Benchmark}
LPMNR, 2013.

\bibitem{BoteaHeuristicsVsPlanning}
Adi Botea, Martin Müller, Jonathan Schaeffer
\textit{Using Abstraction for Planning in Sokoban}
Computers and Games, 2002.

\bibitem{Frisch2006PropagationAF}
Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh
\textit{Propagation algorithms for lexicographic ordering constraints}
Artificial Intelligence, 2006, Vol. 170, p.803-834.

\bibitem{Dor1999SOKOBANAO}
Dorit Dor, Uri Zwick
\textit{SOKOBAN and other motion planning problems}
Computational Geometry, 1999, Vol. 13, p.215-228.

\bibitem{Froleyks2017UsingAA}
Nils Christian Froleyks and Tom{\'a}{\^s} Balyo
\textit{Using an Algorithm Portfolio to Solve Sokoban}
SOCS, 2017.

\bibitem{PSpaceComplete}
Joseph C. Culberson
\textit{Sokoban is PSPACE-complete}
1997.

\bibitem{SokoRelevanceCuts}
Andreas Junghanns, Jonathan Schaeffer
\textit{Sokoban: Improving the Search with Relevance Cuts}
Journal of Theoretical Computing Science, 1999, Vol. 252, p.1-2.

\bibitem{SymmetryBreaking}
Brahim Hnich, Zeynep Kiziltan, Toby Walsh
\textit{Combining Symmetry Breaking with Other Constraints: Lexicographic Ordering with Sums}
ISAIM, 2004.

\bibitem{WilfongNPhard}
Gordon Wilfong
\textit{Motion planning in the presence of movable obstacles}
Annals of Mathematics and Artificial Intelligence, 1991.
\end{thebibliography}

>>>>

\documentclass{article}
\title{Efficient Planning for Sokoban with Answer Set Programming - A Case Study}
\author{Thomas Verweyen}
\date{\today}
\usepackage{listings}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{color}
\lstset{numbers=left,
literate = {-}{-}1,
basicstyle=\footnotesize,
numbersep=5pt, backgroundcolor=\color{white},
captionpos=b,keywordstyle=\color{blue},
commentstyle=\color{green}}
\begin{document}
\maketitle
\renewcommand*\contentsname{Summary}
\tableofcontents
\newpage
\section*{Abstract}
Sokoban approach with ASP planning $\rightarrow$ PlainEnc, IncEnc, Optimization (?)
\section{Introduction}
%%%%TODO: REWORK THIS
What do i do? I'm working on the Sokoban problem and trying to do my best to find a fast and easy solution to it. I'm not working on any level but just the ones meeting some criteria. I'm picking qualitative and quantitative characteristics and deciding on some values to examine. My encodings are supposed to work better on some of the instances and worse on others. What rules describe those fluctuations? What assumptions can be made? Is there a trend when comparing runtimes to my qualitative and quantitative characteristics? E.g. you would expect the runtime to go up when the instance's size grows or when the amount of boxes in the level grows.
In this thesis I will work on optimizing ASP encodings for the game Sokoban towards different instances. First Ill briefly explain what Sokoban is and how it is played, then Ill give a short introduction on comparable/similar/other scientific studies concerning sokoban. Then Ill go over my work, ... and finally evaluation and Discussion of Results.
\newpage
\section{What is Sokoban}
\subsection{Background}
%% ZU KURZ, MIR FEHLT EIN BISSCHEN LORE (SOKOBAN HEISST WARENHAUSMANN, ENTWICKELT VON.., ...) UND EIN BISSCHEN FORMALITÄT (VGL \cite{Froleyks2017UsingAA})
Sokoban is a logistic planning puzzle game from Japan. Sokoban was apparently created by Hiroyuki Imabayashi in 1981, and published by \textit{Thinking Rabbit} in 1982. Sokoban means warehouse person in japanese. The game has very simple rules but shows to be very complex for humans and for AI. This makes Sokoban a very interesting subject for research regarding planning problems.
%LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
\subsection{Rules}
\begin{figure}[h]
\centering
\includegraphics[scale=0.2]{SokobanLevel1}
\caption{First of the Classical Sokoban Levels}
\label{fig:1}
\end{figure}
Figure \ref{fig:1} illustrates the first Sokoban level of the original set of levels. The game is played on a board that can be modelled as a two-dimensional grid. The squares can either be floor or wall squares. Some floor squares are marked as goals and the same number of floor squares contain boxes. The player pushes boxes around this modelled warehouse, trying to cover the marked goal fields (targets) with one of the boxes. For that the player can move one step at a time into one of the four directions \textit{up}, \textit{down}, \textit{left} and \textit{right}. When the player moves onto a box's position the box is pushed one field into the direction of the move. There can never be multiple boxes on one field, neither can there be a box where there is no field. A \textit{solution} to a level is a sequence of moves that leaves a box on each target.
In many of the levels created to this date, the warehouse has a lot of small passageways and very intricate designs. In this thesis I will look at more open level designs, e.g. rectangular levels.
Sokoban is a very difficult planning problem for multiple reasons. There are a lot of possibilities to create a deadlock (a gamestate from which no solution can be found), the branching factor is large and the optimal solution can be very long \cite{BoteaHeuristicsVsPlanning}.

@@

\newpage

@@

\subsubsection{Incremental Encoding}
\paragraph*{•}
\lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
The incremental encoding is similar to the plain encoding. The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t (STEP ALSO IS THOUGH). This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable.

\subsubsection{Advanced Encoding}
\paragraph*{•}
\lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
\subsection{Evaluation}
\paragraph*{•}
\subsection{Experiments and Results}
\paragraph*{•}
\newpage
\section{Discussion of Results}
\paragraph*{•}
compare chapter State of the Art with chapter My Work \cite{LifschitzASP}

\begin{thebibliography}{}
\bibitem{LifschitzASP}
Vladimir Lifschitz,
\textit{Answer Set Planning},
ICLP, 1999.

\bibitem{Lifschitz2002AnswerSP}
Vladimir Lifschitz,
\textit{Answer set programming and plan generation},
Artificial Intelligence, 2002, Vol. 138, p.39-54.

\bibitem{LifschitzWhatisASP}
Vladimir Lifschitz,
\textit{What is Answer Set Programming?},
AAAI, 2008.

\bibitem{Gebser2013RicochetRA}
Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer,
\textit{Ricochet Robots: A Transverse ASP Benchmark},
LPMNR, 2013.

\bibitem{BoteaHeuristicsVsPlanning}
Adi Botea, Martin Müller, Jonathan Schaeffer,
\textit{Using Abstraction for Planning in Sokoban},
Computers and Games, 2002.

\bibitem{Frisch2006PropagationAF}
Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh,
\textit{Propagation algorithms for lexicographic ordering constraints},
Artificial Intelligence, 2006, Vol. 170, p.803-834.

\bibitem{Dor1999SOKOBANAO}
Dorit Dor, Uri Zwick,
\textit{SOKOBAN and other motion planning problems},
Computational Geometry, 1999, Vol. 13, p.215-228.

\bibitem{Froleyks2017UsingAA}
Nils Christian Froleyks and Tom{\'a}{\^s} Balyo,
\textit{Using an Algorithm Portfolio to Solve Sokoban},
SOCS, 2017.

\bibitem{PSpaceComplete}
Joseph C. Culberson,
\textit{Sokoban is PSPACE-complete},
1997.

\bibitem{SokoRelevanceCuts}
Andreas Junghanns, Jonathan Schaeffer,
\textit{Sokoban: Improving the Search with Relevance Cuts},
Journal of Theoretical Computing Science, 1999, Vol. 252, p.1-2.

\bibitem{SymmetryBreaking}
Brahim Hnich, Zeynep Kiziltan, Toby Walsh,
\textit{Combining Symmetry Breaking with Other Constraints: Lexicographic Ordering with Sums},
ISAIM, 2004.

\bibitem{WilfongNPhard}
Gordon Wilfong,
\textit{Motion planning in the presence of movable obstacles},
Annals of Mathematics and Artificial Intelligence, 1991.
\end{thebibliography}
